sentiment-analysis

### 2016-08-08 eda/sentiment-analysis.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(jsonlite)
library(stringr)
library(ggplot2)
library(dplyr)
library(tidytext)
library(tm)
library(ggrepel)
source("../functions/PercentAccuracy.R")
```

@@

Too slow. I'll try the library __tidytext__. This library also let's me generate valence scores using the AFINN library without having to write an unnecessarily long function.

```{r, eval = T}
# unnest_tokens splits column into tokens using tokenizers package
# splits "text" column into individual words to analyze sentiment for further
# exploration
reviews_text <- reviews %>% 
  select(business_id, review_id, stars, text) %>%
  unnest_tokens(word, text) %>% 
  anti_join(stop_words) %>% 
  arrange(review_id)

@@

# calculate mean for each group of stars
# Tidied csv: review_id, star rating, average sentiment score
review_scores <- reviews_text %>% 
  inner_join(valence_score, by = "word") %>% 
  group_by(review_id, stars) %>% 
  summarise(sentiment = mean(score))
```

@@

Positive correlation exists! Some inaccuracy (outliers on boxplots). Find out most common positive and negative words in reviews, creating a summary of the counts. I filtered for greater than 500 reviews and 10 business so they're not unique words to a local business or rare words with strange sentiments.
```{r}
# create summary of words and their average sentiment score in each unique
# review and filter for >= 500 reviews and >= 10 businesses
# Tidied csv: word, # of businesses word appears, # of reviews word appears,
# average stars of word
review_words <- reviews_text %>% 
  count(review_id, business_id, stars, word) %>% 
  ungroup()
summaries <- review_words %>% 
  group_by(word) %>% 
  summarise(business = n_distinct(business_id),
            reviews = n(),
            average_star = mean(stars)) %>% 
  ungroup() %>% 
  filter(reviews >= 500, business >= 10) %>% 
  inner_join(valence_score, by = "word")
head(summaries, n = 10)
```

@@

```{r}
positive <- summaries %>% 
  filter(score > 0)
negative <- summaries %>% 
  filter(score < 0)
percent_pos <- PercentAccuracy(positive)
percent_neg <- 1 - PercentAccuracy(negative)
print(str_c("Positivity Accuracy: ", percent_pos))
print(str_c("Negativity Accuracy: ", percent_neg))
```
64.23% of positive words lie above the mean average rating! There's definitely a positive correlation between word positivity in text reviews and Yelp business star ratings. Word negativity is even better at 87.24% accuracy of negative words lying below the mean average rating.

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(jsonlite)
library(stringr)
library(ggplot2)
library(dplyr)
library(tidytext)
library(tm)
library(purrr)
source("../functions/PercentAccuracy.R")
source("../functions/MeanStar.R")
```

@@

Too slow. I'll try the library __tidytext__. This library also let's me generate valence scores using the AFINN library without mhaving to write an unnecessarily long function.

```{r, eval = T}
# unnest_tokens splits column into tokens using tokenizers package
# splits "text" column into individual words to nalyze sentiment for further
# exploration
reviews_text <- reviews %>% 
  select(business_id, review_id, stars, text) %>%
  unnest_tokens(word, text) %>% 
  anti_join(stop_words) %>% 
  arrange(review_id)

@@

# calculate mean for each review
# Tidied: review_id, star rating, average sentiment score
review_scores <- reviews_text %>% 
  inner_join(valence_score, by = "word") %>% 
  group_by(review_id, stars) %>% 
  summarise(sentiment = mean(score))
write_csv(review_scores, "../clean_data/review_scores.csv")
```

@@

Positive correlation exists! Some inaccuracy (outliers on boxplots). Find out most common positive and negative words in reviews, creating a summary of the counts. I filtered for greater than 500 reviews and 10 business so they're not unique words to a local business or rare words with strange sentiments.
```{r}
# create summary of words and their average sentiment score in each unique
# review and filter for >= 500 reviews and >= 10 businesses
# Tidied: word, # of businesses word appears, # of reviews word appears,
# average stars of word
review_words <- reviews_text %>% 
  count(review_id, business_id, stars, word) %>% 
  ungroup()
summaries <- review_words %>% 
  group_by(word) %>% 
  summarise(business = n_distinct(business_id),
            reviews = n(),
            average_star = mean(stars)) %>% 
  ungroup() %>% 
  filter(reviews >= 500, business >= 10) %>% 
  inner_join(valence_score, by = "word")
head(summaries, n = 10)
write_csv(summaries, "../clean_data/summaries.csv")
```

@@

```{r}
positive <- summaries %>% 
  filter(score > 0)
negative <- summaries %>% 
  filter(score < 0)
percent_pos <- PercentAccuracy(positive)
percent_neg <- 1 - PercentAccuracy(negative)
print(str_c("Positivity Accuracy: ", percent_pos))
print(str_c("Negativity Accuracy: ", percent_neg))
```
63.71% of positive words lie above the mean average rating! There's definitely a positive correlation between word positivity in text reviews and Yelp business star ratings. Word negativity is even better at 87.76% accuracy of negative words lying below the mean average rating.

@@

```{r}
data.frame(
  x = factor(-5:5), 
  y = unlist(-5:5 %>% map(MeanStar))) %>% 
  filter(!is.nan(y)) %>% 
  ggplot(aes(x = x, y = y, fill = x)) +
  geom_bar(stat = "identity") +
  labs(x = "Valence Score",
       y = "Average Star Rating",
       fill = "Valence Score")
```
